Constraints on the mass and occupation fraction of dark matter halos hosting Lyman-alpha emitters at $z\sim 3$

### 2013-07-09 paper/ms.tex
included all the comments from Peter Laursen

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$, by comparing the number density and the angular
correlation function between mock and observed fields. The mock fields
are constructed in a large cosmological N-body simulation using a
model where a dark matter halo can only host one LAE with a
probability $f_{\rm occ}$ if its mass is within the mass range
delimited by two threshold values  $M_{\rm min}$ and $M_{\rm max}$. We
find three different families of parameters based on their mass range
$\Delta M \equiv \log_{10}M_{\rm max} - \log_{10} M_{\rm min}$ dex and occupation
fraction $f_{\rm occ}$.  From this we  conclude that the number
density and spatial clustering information available from observations
is not sufficient to derive a unique and narrow set of values for the
parameters in the model.  However, there is a dominant familiy composed
by models with a narrow mass range $\Delta M<1.0$ dex, a low
occupation fraction $f_{\rm occ}\leq 0.3$ and a maximum mass $M_{\rm
  max}< 10^{12}$\hMsun.  This suggests that the most massive
dark matter halos at that epoch do not host the brightest LAEs, also
giving support to some observational evidence indicating that only a small
fraction of star forming galaxies can be actually detected as LAEs. We
provide tables and mock catalogs for the successful models.  
\end{abstract}

@@

Lyman-$\alpha$ emitting galaxies (LAEs) have become in the last decade a 
central topic in studies of structure formation in the Universe. They 
are helpful in a diverse range of fields. LAEs can
be used as probes of reionization \citep{Dijkstra11}, tracers of
large scale structure \citep{Koehler2007}, 
signposts for low metallicity stellar populations and markers of the
the galaxy formation process through cosmic history \citep{ForeroRomero2012}.

@@

In our model a dark matter halo can only host one or zero LAE. 
There are three parameters in the model that decide whether a halo can host a
galaxy or not: the lower and upper bounds for the mass range $M_{\rm min}<
M_{\rm h} < M_{\rm max}$ where LAEs reside and the fraction $f_{\rm
  occ}$ of such halos that effectively host a LAE. We do not assign a
luminosity to each LAE. We are primarily interested in
constraining the halo mass range hosting detectable LAEs under the
conditions defined by \cite{Yamada2012}. In what follows we will describe by
the letter ${\mathcal M}$ a model defined by an specific choice of the
three scalar parameters $M_{\rm min}$, $M_{\rm  max}$ y $f_{\rm occ}$.

@@

We begin by considering that  a model ${\mathcal M}$ that has at least
one (1) mock survey (out of 15) consistent with the observed
distribution of LAE number densities has viable parameters that
deserve to be considered for further analysis. Later on we impose
harder constraints to reduce the number of models by asking that all
the 15 mocks to be consistent with observations.

@@

@@

\subsection{Three halo mass ranges}

\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig2_match_P5.pdf}
\vspace{5mm}
\includegraphics[width=0.49\linewidth,angle=0]{./plots/Fig3_match_P5.pdf}\\
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig2_random_P5.pdf}
\hspace{5mm}
\includegraphics[width=0.49\linewidth,angle=0]{./plots/Fig3_random_P5.pdf}\\
\end{center} 
\caption{$M_{\rm min}$-$M_{\rm max}$ (left) and $M_{\rm
    min}-f_{\rm occ}$ (right) planes for all models with
  $P>0.05$ in two different ways used to construct the mock
  surveys. The color code corresponds to the number of mock surveys
  that are found to be compatible with observations in terms of the KS
  test with $P>0.05$. Only regions of parameter space with at least
  one (1) consistent mock survey are included. \label{fig:landscape}}   
\end{figure*}

@@

The upper (lower) panels correspond to the {\texttt match} ({\texttt
  random}) method to build the mock surveys from individual
fields. The plot shows number of mock surveys consistent
with observations. There are between $550$ to $600$ models out of the
original $9000$ models that have at least one (1) mock survey
consistent with observations.

@@

For each model ${\mathcal M}$ there are 15 different mock surveys. In the
previous section we presented the models that had at least one (1)
mock survey with $P>0.05$.

@@

Matching the galaxy surface number density statistics sets the median
mass of all successful models in the range $(10^{10}-10^{12})$
\hMsun. Including more strict criteria on the number of mock surveys
that must be consistent with those observations and additional
information from the angular correlation function greatly reduces the
number of possible models. We end up with $\sim 50$ models out of the
initial 90000 possible combination of parameters.

We find that to make a physical interpretation of these models it is
useful to do it in terms of the size of halo mass range. We now use the
variable $\Delta M=\log_{10}M_{\rm max} - \log_{10}M_{\rm  min}$, together
with the escape fraction $f_{\rm occ}$ to help us build a classification
of all the successful models into three families:

@@

At the low mass end, such cut can be readily interpreted in terms of the
minimal halo star formation rate needed to produce the necessary \lya
luminosity to be above a given detection threshold.  However, under
the reasonable assumption of star formation rate increasing with halo
mass, the cut at higher halo masses requires a different explanation.

One possible interpretation can be made in terms of a decreasing escape
fraction of \ly radiation in massive systems. There are detailed models for
radiative transfer that support the idea that massive galaxies with
higher metallicities have larger dust contents than lower mass
systems, which due to the resonant nature of the \ly line are enough
to produce high absorption of \ly photons but not of continuum or
other non-resonant lines.

@@

@@

\section*{Acknowledgments} 
J.E.F-R thanks the hospitality of Changbom Park and the Korea
Institute for Advanced Study where the first full draft of this paper
was completed.

>>>>

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$, by comparing the number density and the angular
correlation function between mock and observed fields. The mock fields
are constructed in a large cosmological N-body simulation using a
model where a dark matter halo can only host one LAE with a
probability $f_{\rm occ}$  if its mass is within the mass range
delimited by two threshold values  $M_{\rm min}$ and $M_{\rm max}$. We
define three different families of parameters consistent with
observations based on their mass range
$\Delta M \equiv \log_{10}M_{\rm max} - \log_{10} M_{\rm min}$ dex and occupation
fraction $f_{\rm occ}$.  From this we conclude that the number
density and spatial clustering information available from observations
is not sufficient to derive a unique and narrow set of values for the
parameters in the model. In particular, it is always possible to find a 
a model with any desired occupation fraction in the range $0.1\leq
f_{\rm occ}\leq 1.0$.  However, there is a dominant family composed by
models with a narrow mass range $\Delta M<1.0$ dex, a low occupation fraction
$f_{\rm occ}\leq 0.3$ and a maximum mass $M_{\rm   max}<
10^{12}$\hMsun.  This suggests that the most massive dark matter halos
at that epoch do not host the brightest LAEs, also giving support to
some observational evidence indicating that only a small fraction of
star forming galaxies can be actually detected as LAEs. We provide
tables and mock catalogs for the successful models.   
\end{abstract}

@@

Lyman-$\alpha$ emitting galaxies (LAEs) have become in the last decade
a  central topic in studies of structure formation in the Universe. They 
are helpful in a diverse range of fields. LAEs can be used as probes
of reionization \citep{Dijkstra11}, tracers of large scale structure
\citep{Koehler2007},  signposts for low metallicity stellar
populations and markers of thethe galaxy formation process through
cosmic history \citep{ForeroRomero2012}.

@@

In our model a dark matter halo can only host one or zero LAE. 
There are three parameters in the model that decide whether a halo can host a
galaxy or not: the lower and upper bounds for the mass range $M_{\rm min}<
M_{\rm h} < M_{\rm max}$ where LAEs reside and the fraction $f_{\rm
  occ}$ of such halos that host a \emph{detectable LAE}. The reader
must keep in mind that the physical interpreation of the occupation
fraction $f_{\rm occ}$ convolves two things: the presence of a star
forming galaxy in a halos and its detectability as a LAE. We do not assign a
luminosity to each LAE. We are primarily interested in
constraining the halo mass range hosting detectable LAEs under the
conditions defined by \cite{Yamada2012}. In what follows we will describe by
the letter ${\mathcal M}$ a model defined by an specific choice of the
three scalar parameters $M_{\rm min}$, $M_{\rm  max}$ y $f_{\rm occ}$.

@@

We begin by considering that  a model ${\mathcal M}$ that has at least
one mock survey (out of 15) consistent with the observed
distribution of LAE number densities has viable parameters that
deserve to be considered for further analysis. Later on we impose
harder constraints to reduce the number of models by asking that all
the 15 mocks to be consistent with observations.

@@

The main purpose of this section is to show how different
observational constraints narrow down the parameters space of allowed
models. In each one of the four sub-sections we add a new
constraint. First we consider the average halo number density, then we
consider the surface density distributions for all the mock fields in
a given mock survey. From this result we move to narrower constraints
by requiring a high number of mock surveys with consistent surface
density distributions to finally consider the impact of the angular
correlation function.

@@

\subsection{Surface density distributions}

\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig2_match_P5.pdf}
\vspace{5mm}
\includegraphics[width=0.49\linewidth,angle=0]{./plots/Fig3_match_P5.pdf}\\
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig2_random_P5.pdf}
\hspace{5mm}
\includegraphics[width=0.49\linewidth,angle=0]{./plots/Fig3_random_P5.pdf}\\
\end{center} 
\caption{$M_{\rm min}$-$M_{\rm max}$ (left) and $M_{\rm
    min}-f_{\rm occ}$ (right) planes for all models with
  $P>0.05$ in two different ways used to construct the mock
  surveys. The color code corresponds to the number of mock surveys
  that are found to be compatible with observations in terms of the KS
  test with $P>0.05$. Only regions of parameter space with at least
  one consistent mock survey are included. \label{fig:landscape}}   
\end{figure*}

@@

The upper (lower) panels correspond to the {\texttt match} ({\texttt
  random}) method to build the mock surveys from individual
fields. The plot shows number of mock surveys consistent
with observations. There are between $550$ to $600$ models out of the
original $9000$ models that have at least one mock survey
consistent with observations.

@@

For each model ${\mathcal M}$ there are 15 different mock surveys. In the
previous section we presented the models that had at least one
mock survey with $P>0.05$.

@@

Matching the galaxy surface number density statistics sets the median
mass of all successful models in the range $(10^{10}-10^{12})$
\hMsun. Including more strict criteria on the number of mock surveys
that must be consistent with those observations and additional
information from the angular correlation function greatly reduces the
number of possible models. We end up with $\sim 50$ models out of the
initial 9000 possible combination of parameters.

We find that to make a physical interpretation of these models it is
useful to do it in terms of the size of halo mass range. We now use the
variable $\Delta M=\log_{10}M_{\rm max} - \log_{10}M_{\rm  min}$, together
with the occupation fraction $f_{\rm occ}$ to help us build a classification
of all the successful models into three families:

@@

At the low mass end, such cut can be readily interpreted in terms of the
minimal halo star formation rate needed to produce the necessary \lya
luminosity to be above a given detection threshold.  However, under
the reasonable assumption of star formation rate increasing with halo
mass, the cut at higher halo masses requires a different
explanation. There are two different physical scenarios to explain
this.

One possible interpretation can be made in terms of a decreasing escape
fraction of \ly radiation in massive systems. There are detailed models for
radiative transfer that support the idea that massive galaxies with
higher metallicities have larger dust contents and less concentrated
ISM than lower mass systems, which due to the resonant nature of the
\ly line are enough to produce high absorption of \ly photons but not of continuum or
other non-resonant lines \citep{ForeroRomero2011,Laursen2009}.

A second scenario is that larger systems have more extended gaseous
envelopes which, due to resonance effects of the \ly line, leading to
a low surface brightness and a broader line, making these systems less
observable in narrow band filter surveys \citep{Zheng2010}.

@@

\citep{Ouchi2010} prefers a very lowe duty cycle $0.008\pm 0.03$.

@@

\section*{Acknowledgments} 
J.E.F-R thanks the hospitality of Changbom Park and the Korea
Institute for Advanced Study where the first full draft of this paper
was completed. J.E.F-R also thanks Peter Laursen for helpful comments
on the physical interpretation and the presentation of our results.
